---
title: "Data Management for cBioPortal"
author: "Kemmapon Chumchuen"
date: "5/25/2021"
output: html_document
---

```{r, message=F, warning=F}
library(epicalc)
library(stringr)
library(BSL)
library(readxl)
library(Hmisc)
zap()
```


Load cleaned dataset (should be the data from Data prep.R)

```{r}
data <- read_excel("Human.1.0.xlsx", skip = 11, sheet = 1)
```


Check for column names

```{r}
colnames(data)
#str(data)
```

Select variables to import 

```{r}
select <- c(1:3,5:6,8:9,15,17,21,23)
data <- data[,select]
colnames(data)
```


Identify PATIENT_ID and SAMPLE_ID(required) columns and rename

```{r, echo=F}
colnames(data)[2] <- "PATIENT_ID" 
colnames(data)[1] <- "SAMPLE_ID"
colnames(data)
data <- cbind(data[,2],data[,-2])
colnames(data)
```

Change other column names to uppercase

```{r}
colnames(data)[] <- toupper(colnames(data))
colnames(data)
```

In case _ is needed

```{r}
#colnames(data) <- gsub(" ", "_", colnames(data))
```

Change the column names to match the format of cBioPortal survival plot

```{r}
#colnames(data)[7:8] <- c("DS_STATUS","DS_YRS")
#colnames(data)
```

Delete the unwant symbol in the column names (from the original dataset)

```{r}
colnames(data) <- gsub("\\*","",colnames(data))
colnames(data)
```

Creat TNM_STAGE for the trial run

```{r}
tnm <- c("T0","T1","T2","T3","T4","N0","N1","N2","N3","M0","M1")
data$TNM_STAGE <- sample(tnm,116,replace = T)
```



Create metadata, if the data did not come from REDcap or contain no labels,
specify labelled = FALSE

```{r, message=F, warning=F}
meta <- cbio_meta(data = data, labelled = FALSE)
meta
```

In case of no labels, manually add the labels (description) in row 2

```{r}
meta[2,] <- c("#Patient identifer","Sample identifier","Accession number","Sample isolate","Age",
              "Sex","Type of tissue","Disease","Patient's ethnicity","Nationality",
              "Type of sample","TNM staging for breast cancer.")
meta
```



Add meta data to the dataset and export in tab seperated format (patient and sample data)

```{r}
colnames(meta) <- colnames(data)
met.dat <- rbind(meta,data)
pat <- c(1,5,6,9,10)
pat.dat <- met.dat[,pat]
sam <- c(1,2,3,4,7,8,11,12)
sam.dat <- met.dat[,sam]

write.table(pat.dat, file='data_clinical_patient.txt', quote=FALSE, sep='\t', 
            row.names = F, fileEncoding = "UTF-8",col.names = F)

write.table(sam.dat, file='data_clinical_sample.txt', quote=FALSE, sep='\t', 
            row.names = F, fileEncoding = "UTF-8",col.names = F)
```
